---
title: "p8105_hw3_jz3425"
author: "Jibei Zheng"
date: 10/17/2021
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(p8105.datasets)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Load instacart dataset.

```{r}
data("instacart")

count_users = count(instacart, user_id)
```

__Description of the dataset:__

This instacart dataset is an anonymized dataset with `r nrow(instacart)` observations of `r nrow(count_users)` unique users and one order from each of them. Each row represents one product of an order. There are `r ncol(instacart)` variables. Key variables include `user_id` and `product_id`, `add_to_cart_order`: order in which each product was added to cart, `reordered`: if one product has been reordered, `order_dow` and `order_hour_of_day`: day and time an order was placed, and categories information of each product.


For example:

```{r}
#example df
order_1 = instacart %>% 
  filter(order_id == 1)

knitr::kable(order_1, digits = 1)
```

These rows show an order from user `r order_1[1, 5]` of `r nrow(order_1)` products by the order in which they were added to the cart, each with name, aisle and department information. The order was placed on Thursday `r order_1[1, 9]`(in 24 hour system). 4 products were reordered. It had been `r order_1[1, 10]` days since the last order of this user. 

__·How many aisles are there, and which aisles are the most items ordered from?__

```{r}
count_aisle = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))
count_aisle
```

There are `r nrow(count_aisle)` aisles, and fresh vegetables are the most items ordered from.

__·Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.__

```{r}
count_aisle %>% 
  filter(n_obs > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n_obs)) %>% 
  ggplot(aes(x = n_obs, y = aisle)) +
  geom_col() +
  labs(
    title = "Aisle Plot",
    x = "Number of items ordered",
    y = "Aisle"
  ) +
  scale_x_continuous(breaks = c(10000, 30000, 50000, 80000, 150000))
```

There are 39 aisles with more than 10000 items ordered, among which fresh vegetables and fresh fruits are the most popular with more than 150000 orders.

__·Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.__

```{r, message = FALSE}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_items = n()) %>% 
  filter(min_rank(desc(n_items)) < 4) %>% 
  arrange(desc(n_items), .by_group = TRUE) %>% 
  knitr::kable(digits = 1)
```

Among these three aisles, the most popular one is packaged vegetables fruits, each product with thousands of orders; the least popular one is dog food care, each product with only less than 30 orders, probably because they are less frequently needed.

```{r, message = FALSE}
#generate a plot
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_items = n()) %>% 
  filter(min_rank(desc(n_items)) < 4) %>% 
  mutate(product_name = fct_reorder(product_name, n_items)) %>% 
  ggplot(aes(y = product_name, x = n_items, fill = aisle)) +
  geom_col() +
  labs(
    title = "Most Popular Items",
    x = "Number of items ordered",
    y = "Item"
  ) +
  scale_x_continuous(
    trans = "sqrt",
    breaks = c(30, 400, 5000, 10000)
  ) +
  facet_grid(aisle ~ ., scales = "free_y")
```

__·Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).__

```{r, message = FALSE}
dow = c("monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday")

instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  mutate(order_dow = dow[order_dow + 1]) %>% 
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hour"
  ) %>% 
  knitr::kable(digits = 1)
```

On most of the days, people buy coffee ice creams a little later than pink lady apples, conforming to the finding that healthier food tends to be purchased earlier in the day.

## Problem 2

Load BRFSS SMART 2010 data and clean it.

```{r, warning = FALSE}
data("brfss_smart2010")

brfss_df = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  rename(state = locationabbr, county = locationdesc) %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) 
```

__·In 2002, which states were observed at 7 or more locations? What about in 2010?__

```{r}
brfss_df %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarize(n_obs = n() / 5) %>% 
  filter(n_obs >= 7) %>% 
  knitr::kable(digits = 1)
```

There were 6 states listed above which observed at 7 or more locations.

```{r}
brfss_df %>% 
  filter(year == 2010) %>% 
  group_by(state) %>% 
  summarize(n_obs = n() / 5) %>% 
  filter(n_obs >= 7) %>% 
  knitr::kable(digits = 1)
```

There were 14 states listed above which observed at 7 or more locations, 8 more than 2002.

__Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state.__

```{r, fig.height=10, message = FALSE, warning = FALSE}
brfss_df %>% 
  filter(response == "Excellent") %>% 
  group_by(year, state) %>% 
  summarize(average_data_value = mean(data_value)) %>% 
  ggplot(aes(x = year, y = average_data_value, color = state)) +
  geom_line()
```

Maybe it's a bad idea to put all the lines in one plot.