---
title: "p8105_hw3_jz3425"
author: "Jibei Zheng"
date: 10/17/2021
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(p8105.datasets)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Load instacart dataset.

```{r}
data("instacart")

count_users = count(instacart, user_id)
```

__Description of the dataset:__

This instacart dataset is an anonymized dataset with `r nrow(instacart)` observations of `r nrow(count_users)` unique users and one order from each of them. Each row represents one product of an order. There are `r ncol(instacart)` variables. Key variables include `user_id` and `product_id`, `add_to_cart_order`: order in which each product was added to cart, `reordered`: if one product has been reordered, `order_dow` and `order_hour_of_day`: day and time an order was placed, and categories information of each product.


For example:

```{r}
#example df
order_1 = instacart %>% 
  filter(order_id == 1)

knitr::kable(order_1, digits = 1)
```

These rows show an order from user `r order_1[1, 5]` of `r nrow(order_1)` products by the order in which they were added to the cart, each with name, aisle and department information. The order was placed on Thursday `r order_1[1, 9]`(in 24 hour system). 4 products were reordered. It had been `r order_1[1, 10]` days since the last order of this user. 

__·How many aisles are there, and which aisles are the most items ordered from?__

```{r}
count_aisle = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))
count_aisle
```

There are `r nrow(count_aisle)` aisles, and fresh vegetables are the most items ordered from.

__·Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.__

```{r}
count_aisle %>% 
  filter(n_obs > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n_obs)) %>% 
  ggplot(aes(x = n_obs, y = aisle)) +
  geom_col() +
  labs(
    title = "Aisle Plot",
    x = "Number of items ordered",
    y = "Aisle"
  ) +
  scale_x_continuous(breaks = c(10000, 30000, 50000, 80000, 150000))
```

There are 39 aisles with more than 10000 items ordered, among which fresh vegetables and fresh fruits are the most popular with more than 150000 orders.

__·Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.__

```{r, message = FALSE}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_items = n()) %>% 
  filter(min_rank(desc(n_items)) < 4) %>% 
  arrange(desc(n_items), .by_group = TRUE) %>% 
  knitr::kable(digits = 1)
```

Among these three aisles, the most popular one is packaged vegetables fruits, each product with thousands of orders; the least popular one is dog food care, each product with only less than 30 orders, probably because they are less frequently needed.

```{r, message = FALSE}
#generate a plot
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_items = n()) %>% 
  filter(min_rank(desc(n_items)) < 4) %>% 
  mutate(product_name = fct_reorder(product_name, n_items)) %>% 
  ggplot(aes(y = product_name, x = n_items, fill = aisle)) +
  geom_col() +
  labs(
    title = "Most Popular Items",
    x = "Number of items ordered",
    y = "Item"
  ) +
  scale_x_continuous(
    trans = "sqrt",
    breaks = c(30, 400, 5000, 10000)
  ) +
  facet_grid(aisle ~ ., scales = "free_y")
```

__·Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).__

```{r, message = FALSE}
dow = c("monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday")

instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  mutate(order_dow = dow[order_dow + 1]) %>% 
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hour"
  ) %>% 
  knitr::kable(digits = 1)
```

On most of the days, people buy coffee ice creams a little later than pink lady apples, conforming to the finding that healthier food tends to be purchased earlier in the day.

## Problem 2

Load BRFSS SMART 2010 data and clean it.

```{r, warning = FALSE}
data("brfss_smart2010")

brfss_df = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  rename(state = locationabbr, county = locationdesc) %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) 
```

__·In 2002, which states were observed at 7 or more locations? What about in 2010?__

```{r}
brfss_df %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarize(n_obs = n() / 5) %>% 
  filter(n_obs >= 7) %>% 
  knitr::kable(digits = 1)
```

There were 6 states listed above which observed at 7 or more locations.

```{r}
brfss_df %>% 
  filter(year == 2010) %>% 
  group_by(state) %>% 
  summarize(n_obs = n() / 5) %>% 
  filter(n_obs >= 7) %>% 
  knitr::kable(digits = 1)
```

There were 14 states listed above which observed at 7 or more locations, 8 more than 2002.

__Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state.__

```{r, fig.height=10, message = FALSE, warning = FALSE}
excellent_df =
  brfss_df %>% 
  filter(response == "Excellent") %>% 
  group_by(year, state) %>% 
  summarize(average_data_value = mean(data_value))

excellent_df

excellent_df %>% 
  ggplot(aes(x = year, y = average_data_value, color = state)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Excellenet response over year"
  )
```

Maybe it's a bad idea to put all the lines in one plot.

__Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses(“Poor” to “Excellent”) among locations in NY State.__

```{r, fig.height=10}
brfss_df %>% 
  filter(state == "NY") %>% 
  filter(year %in% c(2010, 2006)) %>% 
  select(year, county, response, data_value) %>% 
  ggplot(aes(x = response, y = data_value, fill = county, group = 1)) +
  geom_col() +
  facet_grid(county ~ year) +
  labs(
    title = "Distribution of responses in NY State"
  )
```

## Problem 3

Load and tidy accel_data.

```{r}
accel_df = 
  read.csv(file = "./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    day_type = case_when(
      day == "Saturday" | day == "Sunday" ~ "weekend",
      TRUE ~ "weekday"
    ),
    day = str_to_lower(day),
  ) %>% 
  rename(week_id = week, day_of_week = day) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute_id",
    names_prefix = "activity_",
    values_to = "activity"
  )

accel_df
```

This accelerometers dataset is from contains `r nrow(accel_df)` observations of activity counts for each minute of a 24-hour dayx starting at midnight(1440 per day) within 5 weeks(35 days) of a 63 year-old male with BMI 25. There are `r ncol(accel_df)` variables including week id, day id, minute id and most important, activity. Not very sure why the day of week are in weird order, but I decide to follow the order of day_id first, and I will reorder them chronologically in next steps.

__Aggregate across minutes to create a total activity variable for each day, and create a table showing these totals.__

```{r, message = FALSE}
#table in day_id order
total_activity_day_a =
accel_df %>% 
  group_by(day_id, week_id, day_of_week) %>% 
  summarize(total_activity = sum(activity))

knitr::kable(total_activity_day_a, digits = 1)

#plot in day_id order
total_activity_day_a %>% 
  ggplot(aes(x = day_id, y = total_activity)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Total activity plot in day_id order",
    x = "Day",
    y = "Total Activity"
  ) +
  scale_x_continuous(
    breaks = c(7, 14, 21, 28, 35)
  )
  
#table in chronological order
total_activity_day_b =
accel_df %>% 
  mutate(day_of_week = factor(day_of_week, levels = c("monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday"))) %>% 
  arrange(week_id, day_of_week) %>% 
  group_by(week_id, day_of_week) %>% 
  summarize(total_activity = sum(activity))

knitr::kable(total_activity_day_b, digits = 1)

#plot in chronological order
total_activity_day_b %>% 
  ungroup() %>% 
  mutate(day_id = c(1:35)) %>% 
  ggplot(aes(y = total_activity, x = day_id)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Total activity plot in chronological order",
    x = "Day",
    y = "Total Activity"
  ) +
  scale_x_continuous(
    breaks = c(7, 14, 21, 28, 35)
  )
```

Both plots do not show apparent trends.

__Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week.__

```{r, message = FALSE}
accel_df %>% 
  mutate(
    minute_id = as.numeric(minute_id),
    hour_id = floor((minute_id - 1) / 60)
  ) %>% 
  group_by(day_of_week, hour_id) %>% 
  summarize(hour_activity = sum(activity) / 5) %>% 
  ggplot(aes(x = hour_id, y = hour_activity, color = day_of_week)) +
  geom_point() +
  geom_line() +
  labs(
    title = "24-hour activity plot",
    x = "hour of day",
    y = "hour activity"
  ) +
  scale_x_continuous(
    breaks = c(0, 4, 8, 12, 16, 20, 24),
    labels = c("12am", "4am", "8am", "12pm", "4pm", "8pm", "12am")
  )
```

This is a plot taking the mean of 24-hour activity time of every Monday to Sunday within 5 weeks. This man has a rather regular schedule. We can see that 0-5 a.m. looks like sleeping time, and after 5 a.m. the man starts to move; after 8 p.m. activity starts to reduce. Among day of week, he seems to have high activity at Friday and Monday nights, and low activity at Sunday and Tuesday nights. On Sunday morning he has high activity, maybe having some routines.
